Efficient pedigree recording for fast population genetics simulation

### 2017-11-18 forwards_paper.tex
Add formatting comment

Recovering the sequence of trees from this information is straightforward:
each point along the genome that the tree topology changes
is accompanied by the end of some \emph{edges} and the beginning of others.
Since each \emph{edge} records the genomic interval 
over which a given node inherits from a particular ancestor,
to construct the tree at a certain point in the genome
we need only retrieve all edges overlapping that point
and construct the corresponding tree.
To modify the tree to reflect the genealogy at a nearby location,
we need only remove those edges whose intervals do not overlap that location,
and add those new edges whose intervals do.
Incidentally, this property that edges naturally encode \emph{differences}
between nearby trees (e.g., as ``subtree prune and regraft'' moves)
allows for efficient algorithms that take advantage
of the highly correlated nature of nearby trees.

@@

The tables API is primarily designed to facilitate efficient interchange of
data between programs or between different modules of the same program.
Following the current best-practises [citations: apache arrow, etc] data is stored
in a columnar format.
The principal advantage of this for our purposes is that it allows for very efficient
interchange of large amounts of numerical data. In principle, this enables
zero-copy semantics, where a data consumer can read the information directly
from the memory of a producer without incurring the overhead of a copy
[citation?] Our implementation uses the numpy C API [citation] to efficiently copy
data from Python into the low-level C library used to manipulate
tree sequences.
Thus, by using a simple numerical
encoding of tree topologies and contiguous arrays of data to store this
encoding, we can achieve data transfer rates that would be impossible under
any text-based approach while retaining excellent portability.

>>>>

Recovering the sequence of trees from this information is straightforward:
each point along the genome that the tree topology changes
is accompanied by the end of some \emph{edges} and the beginning of others.
Since each \emph{edge} records the genomic interval 
over which a given node inherits from a particular ancestor,
to construct the tree at a certain point in the genome
we need only retrieve all edges overlapping that point
and construct the corresponding tree.
To modify the tree to reflect the genealogy at a nearby location,
we need only remove those edges whose intervals do not overlap that location,
and add those new edges whose intervals do.
Incidentally, this property that edges naturally encode \emph{differences}
between nearby trees (e.g., as ``subtree prune and regraft'' moves)
allows for efficient algorithms that take advantage
of the highly correlated nature of nearby trees~\citep{kelleher2016efficient}.

@@

The Tables API is primarily designed to facilitate efficient interchange of
data between programs or between different modules of the same program.
\jda{should capitalize Tables, or not, as above}
Following the current best-practises [citations: apache arrow, etc] data is stored
in a columnar format.
The principal advantage of this for our purposes is that it allows for very efficient
interchange of large amounts of numerical data. In principle, this enables
zero-copy semantics, where a data consumer can read the information directly
from the memory of a producer without incurring the overhead of a copy
[citation?] Our implementation uses the numpy C API [citation] to efficiently copy
data from Python into the low-level C library used to manipulate
tree sequences.
Thus, by using a simple numerical
encoding of tree topologies and contiguous arrays of data to store this
encoding, we can achieve data transfer rates that would be impossible under
any text-based approach while retaining excellent portability.
